\documentclass{homework}
\author{Tashfeen, Ahmad}
\class{CSCI 6213: Data Science Fundamentals}
\title{Homework 5}
\address{Oklahoma City University, Petree College of Arts \& Sciences, Computer Science}

\renewcommand\P{\mathbf{P}}

\begin{document} \maketitle
You're \textit{not allowed} to use any libraries other than Numpy or Matplotlib.

We start by reminding ourselves of the variables,
\[
  x_i =
  \begin{bmatrix}
    t_1 \\ \vdots \\ t_{p-1} \\ 1
  \end{bmatrix},
  \quad
  X_{(n, p)} =
  \begin{bmatrix}
    x^T_1 \\ \vdots \\ x^T_n
  \end{bmatrix},
  \quad
  \beta_{(p, 1)} =
  \begin{bmatrix}
    \beta_1 \\ \vdots \\ \beta_p
  \end{bmatrix},
  \quad
  Y_{(n, 1)} =
  \begin{bmatrix}
    y_1 \\ \vdots \\ y_n
  \end{bmatrix}
\]
In the previous assignment, we differentiated the error function and minimised the derivative to get $\beta$,
\[
    \beta = (X^TX)^{-1}X^TY \quad\text{when}\quad \frac{\p (X\beta - Y)^T(X\beta - Y)}{\p \beta} = 0.
\]
In this assignment, we'll explore $\beta$ from a more Bayesian point of view.

We start by clearly laying down our assumptions on the data,

\begin{description}
    \item[Linearity] We assume that there is indeed a linear relationship between $x_i$ (a column vector in $\R^p)$ and $y_i \in \R$. This is mathematically expressed as: the expected or mean value of $y_i$ is a linear function of $x_i$,
    \[
        \mu_i = \mathbb{E}[y_i|x_i] = x_{i}^{T}\beta.
    \]
    \item[Variance] We assume that all $y_i$ have the same variance $\sigma^2$.
    \item[Independent \& Identically Distributed (IID)] We assume that all $y_i$ are independent and are from the same type of distribution.
    \item[Normal Distribution] Finally state that all $y_i$ are normally distributed.
    \[
        y_i \sim \mathcal{N}(x_{i}^{T}\beta, \sigma^2_i)
    \]
\end{description}

We can now state the problem of finding the best $\beta$ as finding the \textit{most probable} mean $x_{i}^{T}\beta$ and variance $\sigma^2_i$ given the data we have.
\[
    \max_\beta \P(x_{i}^{T}\beta, \sigma^2_i\mid y_i)
\]
Once more, by the Bayes' Theorem,
\[
    \P(x_{i}^{T}\beta, \sigma^2_i\mid y_i) =  \frac{\P(y_i\mid x_{i}^{T}\beta, \sigma^2_i)\P(x_{i}^{T}\beta, \sigma^2_i)}{\P(y_i)}
\]
Note that we now require/are allowed to incorporate some prior knowledge of $\beta$'s distribution. For this assignment, we'll assume $\beta$ to be normally distributed around 0 with a variance of $\tau_i$.
\[
    \beta_i \sim \mathcal{N}(0, \tau^2_i)
\]
We may now rewrite $\P(x_{i}^{T}\beta, \sigma^2_i\mid y_i)$ in term of data matrices,
\[
    \P(X\beta, \sigma^2_i\mathbb{I}\mid Y) =  \frac{\P(Y \mid X\beta, \sigma^2_i\mathbb{I})\P(X\beta, \sigma^2_i\mathbb{I})}{\P(Y)}
\]

Due to the IID assumption,
\begin{align*}
  \P(Y \mid X\beta, \sigma^2_i\mathbb{I}) &= \prod_{i=1}^{n}\P(y_i|x_{i}^{T}\beta, \sigma^2_i) \\
     & = \prod_{i=1}^{n} \frac{1}{\sigma_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2}  && \text{By the \href{https://en.wikipedia.org/wiki/Normal_distribution}{PDF of Normal Distribution}}
\end{align*}

Finally,
\begin{align*}
  \max_\beta \P(x_{i}^{T}\beta, \sigma^2_i\mid y_i) &= \max_\beta \frac{\P(Y \mid X\beta, \sigma^2_i\mathbb{I})\P(X\beta, \sigma^2_i\mathbb{I})}{\P(Y)} \\
        &= \max_\beta \frac{\prod_{i=1}^{n} \frac{1}{\sigma_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2}\P(X\beta, \sigma^2_i\mathbb{I})}{\P(Y)} \\
        &= \max_\beta \prod_{i=1}^{n} \frac{1}{\sigma_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2}\P(X\beta, \sigma^2_i\mathbb{I}) \\
        &= \max_\beta \prod_{i=1}^{n} \frac{1}{\sigma_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2}\P(\beta) \\
        &= \max_\beta \prod_{i=1}^{n} \frac{1}{\sigma_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2}\frac{1}{\tau_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\tau_i^2}(0 - \beta_i)^2} \\
        &= \max_\beta \prod_{i=1}^{n} \frac{1}{\sigma_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2}\frac{1}{\tau_i\sqrt{2\pi}}\exp\paren{-\frac{1}{2\tau_i^2}\beta_i^2} \\
        &= \max_\beta \prod_{i=1}^{n} \exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2}\exp\paren{-\frac{1}{2\tau_i^2}\beta_i^2} \\
        &= \max_\beta \prod_{i=1}^{n} \exp\paren{-\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2-\frac{1}{2\tau_i^2}\beta_i^2} \\
        &= \min_\beta \prod_{i=1}^{n} \exp\paren{\frac{1}{2\sigma_i^2}(x_{i}^{T}\beta - y_i)^2+\frac{1}{2\tau_i^2}\beta_i^2} \\
        &= \min_\beta \prod_{i=1}^{n} \exp\paren{(x_{i}^{T}\beta - y_i)^2+\frac{2\sigma_i^2}{2\tau_i^2}\beta_i^2} \\
        &= \min_\beta \prod_{i=1}^{n} \exp\paren{(x_{i}^{T}\beta - y_i)^2+\lambda\beta_i^2} && \lambda = \frac{\sigma_i^2}{\tau_i^2} \\
        &= \min_\beta \sum_{i=1}^{n} \ln\exp\paren{(x_{i}^{T}\beta - y_i)^2+\lambda\beta_i^2} \\
        &= \min_\beta \sum_{i=1}^{n} (x_{i}^{T}\beta - y_i)^2+\lambda\beta_i^2 \\
        &= \min_\beta (X\beta - Y)^T(X\beta - Y)+\lambda\beta^T\beta \\
\end{align*}

At this point, in order to minimise the above, we differentiate $(X\beta - Y)^T(X\beta - Y)+\lambda\beta^T\beta$ and set it equal to zero. From the previous assignment, we know that,
\[
    \frac{\p (X\beta - Y)^T(X\beta - Y)}{\p \beta} = 2X^TX\beta - 2X^TY
\]
And,
\[
    \frac{\p \lambda\beta^T\beta}{\p\beta} = \lambda \beta
\]
Therefore, the derivative equal to zero is as follows,
\[
    2X^TX\beta - 2X^TY + \lambda \beta = 0
\]

\question\label{eq} Let $\mathbb{I}$ be a $p$ by $p$ identity matrix. Solve $2X^TX\beta - 2X^TY + \lambda \beta = 0$ for $\beta$ as the new linear model. Show your work.

\question Using the same two columns you picked in the last assignment from Boston house--price data, draw a (appropriately labelled) plot using the equation you got in question \ref{eq}. The plot should graph integers $-10 < \lambda <20$ on the $x$--axis and the model's root mean squared error on the $y$--axis. Use a file called \texttt{ridge.py} for this question.

\question From your graph, what is the best value of the hyper-parameter $\lambda$ that corresponds to the smallest root mean squared error? Give both the best  $-10 < \lambda <20$ and the corresponding error.

\section*{Submission Instructions}

\begin{enumerate}
    \item Submit a PDF that answers the questions and contains all the plots that the assignment asks for.
    \item Submit your \texttt{ridge.py}.
\end{enumerate}
\end{document}
