\documentclass{homework}
% \author{Lastname, Firstname}  % Uncomment with your name
\class{MATH 3503: Tashfeen's Discrete Mathematics}
% \date{Spring 2023}  % Uncomment with your semester
\title{Homework 5}
\address{
  Computer Science, %
  Petree College of Arts \& Sciences, %
  Oklahoma City University
}

\begin{document} \maketitle

\question Please read chapter 8 of Chartrand et al. and write a couple sentences about a topic/example/concept that you found difficult or interesting and why?

\question Each student at a certain university is given a 6-digit code (such as 123789 or 001122).
\begin{enumerate}[label=(\alph*)]
    \item How many different codes are there?
    \item How many codes read the same forward and backward?
    \item How many codes contain only odd digits?
    \item How many codes contain at least one even digit?
\end{enumerate}

\question How many different 10-bit strings begin with 1011 or 0110?

\question A password on a computer system consists of four characters, each of which is either a digit or a letter of the alphabet. Suppose that each password must contain at least one digit and at least one letter. How many different such passwords are there?

\question A total of 70 students who go to football, basketball or hockey games on a regular basis are surveyed as to which of these three events they attend. They responded:

38 students go to football games. \par
38 students go to basketball games. \par
35 students go to hockey games. \par
17 students go to both football and basketball games. \par
15 students go to both football and hockey games. \par
16 students go to both basketball and hockey games. \\
How many go to all three?

\question How many people must be present to guarantee that
\begin{enumerate}[label=(\alph*)]
    \item at least two have the same birthday?
    \item at least two of their birthdays are in the same month?
    \item at least three of their birthdays are in one of the months January, February, March, April or at least four of their birthdays are in one of the remaining months?
\end{enumerate}

\question How many different 8-bit strings contain exactly 5 1s?

\question Show that $\binom{2n}{2} =2\binom{n}{2} + n^2$.

\question A student committee is to consist of 3 seniors and 4 juniors. A total of 6 seniors and 8 juniors have volunteered to serve on the committee. How many committees are possible?

\question A total of 5 seniors, 3 juniors and 4 sophomores have volunteered to serve on a 4-person committee. How many committees are possible if
\begin{enumerate}[label=(\alph*)]
    \item there is no other restriction on membership for the committee?
    \item at least one senior, one junior and one sophomore must serve on the committee?
    \item at least 3 seniors must serve on the committee?
    \item at least one senior must serve on the committee?
\end{enumerate}


\end{document}
