\documentclass{homework}
% \author{Lastname, Firstname}  % Uncomment with your name
\class{MATH 3503: Tashfeen's Discrete Mathematics}
% \date{Spring 2023}  % Uncomment with your semester
\title{Homework 6}
\address{
  Computer Science, %
  Petree College of Arts \& Sciences, %
  Oklahoma City University
}

\begin{document} \maketitle

\textit{Please give your final answer as a number where possible as well as show work for full credit.}

\question Please read chapter 9 of Chartrand et al. and write a couple sentences about a topic/example/concept that you found difficult or interesting and why?

\question{\label{q16}} What is the exact term in $\paren{\sqrt{x} - \frac{1}{\sqrt{x}}}^{10}$ containing $x$?

\question Use the Binomial Theorem to simplify the sum,
\[
    2^{n^2} + \binom{n}{1}2^{n(n-1)} + \binom{n}{2}2^{n(n-2)} + \cdots + \binom{n}{n-1}2^n + 1
\]

\question How many different 8-digit numbers can be obtained by permuting the digits in the number $70,440,704$? (Since each number is an 8-digit number, the first digit cannot be 0.)

\question A professor has 10 identical new pens that he no longer needs. In how many ways can these pens be given to 3 students if
\begin{enumerate}[label=(\alph*)]
    \item there are no other conditions?
    \item every student must receive at least one pen?
    \item every student must receive at least two pens?
    \item every student must receive at least three pens?
\end{enumerate}

\question How many 4-digit numbers are there, the sum of whose digits is 11?

\question A man enters Ben's Bagels to buy a dozen bagels only to learn that even though Ben has a large supply of plain bagels and blueberry bagels, he only has two cinnamon bagels left and no other kinds.
\begin{enumerate}[label=(\alph*)]
    \item Express the number of selections of $r$ bagels $(r \geq 0)$ the man can make as a product of polynomials and/or power series\footnote{The book has a list of common power series in figure 9.8 on page 347.}.
    \item Determine the coefficient of $x^{12}$ of the product in (a). What information does this coefficient provide?
\end{enumerate}
\question Suppose that we are interested in the number of non-negative integer solutions $a, b$ and $c$ of the equation $a + 2b + 3c = r$, where $r \in \N$.
\begin{enumerate}[label=(\alph*)]
    \item What is this number when $r \in \{1, 2, 3, 4\}$?
    \item Use generating functions to describe this number for a general $r \in \N$.
\end{enumerate}
\question A man buys 6 doughnuts, each of which is a plain doughnut, a powdered doughnut or a glazed doughnut. How many possible selections are there if he buys at least one doughnut of each kind?
\begin{enumerate}[label=(\alph*)]
    \item Answer the question above by listing all possible selections in a table.
          % this is just generic table environment in latex.
          % \begin{table}[hbtp]
          %     \centering
          %     \begin{tabular}{|| c | c | c | c ||}
          %         \hline
          %         A & B & C & D \\ \hline
          %         E & F & G & H \\ \hline
          %         I & G & K & L \\ \hline
          %         M & N & O & P \\ \hline
          %         Q & R & S & T \\ \hline
          %         U & V & W & X \\ \hline
          %         Y & Z &   &   \\ \hline
          %     \end{tabular}
          %     \caption{Selections Table.}\label{tab:sel}
          % \end{table}
    \item Answer the question above by determining the coefficient of $x^6$ in a product of polynomials and/or power series.
    \item Answer the question above by computing $\binom{s+t-1}{s}$ for an appropriate choice of $s$ and $t$.
\end{enumerate}
\end{document}
