\documentclass{homework}
\author{Tashfeen, Ahmad}
\class{CSCI 4203: Machine Learning}
% \date{\today}
\title{Homework 2}
\address{Oklahoma City University, Petree College of Arts \& Sciences, Computer Science}

\begin{document} \maketitle

You're only allowed to use the standard Python 3 libraries with the exception of Numpy and Matplotlib.

\question Plot the following trigonometric functions for $-5\pi \leq x \leq 5\pi$.
\[
      f(x) = x\sin^2(x), \quad  g(x) = -x\sin^2(x)
\]
\begin{itemize}
    \item The title of the plot should read ``Trigonometric Functions.''
    \item The $x$--axis label of the plot should read ``$x$--axis.''
    \item The $y$--axis label of the plot should read ``$y$--axis.''
    \item The legends should appropriately denote the different functions by their definition.
    \item The graphs should have smooth continuos curves.
\end{itemize}

\question Download and save the novel
\href{https://www.gutenberg.org/files/20727/20727.txt}{``The Cosmic
  Computer''} by Piper. Read the saved text file into Python and plot
a bar graph of the character frequency (case insensitive).

\begin{itemize}
    \item The title of the plot should read `Character Frequency in ``The Cosmic Computer'' by Piper.'
    \item The $x$--axis label of the plot should read ``Alphabet.''
    \item The $y$--axis label of the plot should read ``Frequency.''
    \item The bars should be sorted in ascending order.
\end{itemize}

\question Using \texttt{matplotlib.pyplot.imshow} and an eight by eight binary Numpy array, plot the chess board pattern.

\begin{itemize}
    \item The $x$--ticks should be letters from A to H, A being at the left.
    \item The $y$--ticks should be numbers from 1 to 8, one being at the bottom.
    \item The title should be ``Chess Board Pattern.''
\end{itemize}

\question Following are the two matrices from the last homework,
\[
  \mathbf{A} = \begin{bmatrix} 1 & 0 & 1 \\   2 & 1 & 1 \\   0 & 1 & 1 \\   1 & 1 & 2 \\   \end{bmatrix}, \quad
  \mathbf{B} = \begin{bmatrix} 1 & 2 & 1 \\   2 & 3 & 1 \\   4 & 2 & 2 \\   \end{bmatrix}
\]
Use Numpy to multiply them and print the product $\mathbf{AB}$. Does it match with the one you carried out by hand?

\question The Gregory series is given as,
\[
f(n) = \sum_{i=1}^{i=n}\frac{4(-1)^{i+1}}{2i-1}=\frac{4}{1}-\frac{4}{3}+\frac{4}{5}-\cdots+\frac{4(-1)^{n+1}}{2n-1}
\]
\begin{enumerate}
    \item Without using any loops \ie using vectorised operations over Numpy arrays, plot the following values,
    \[
        \{(n, f(n)) : 1 \leq n \leq 100\}
    \]
    \item Plot the following error curve,
    \[
        \{(n, (\pi - f(n))^2) : 1 \leq n \leq 100\}
    \]
\end{enumerate}
Label the axis and title each plot appropriately. What is $\lim_{n\ra\infty}f(n)$?
\section*{Submission Instructions}

\begin{enumerate}
    \item Submit a PDF that answers all the questions and includes any plots that the assignment asks for. Circle and/emphasize your final answer wherever possible.
    \item Submit a Python file \eg \texttt{plot.py} that generates all the plots one by one.
\end{enumerate}

\end{document}
